---
title: "Tidyverse"
author: Kyle Thomas
date: March 22, 2005
output: html_notebook
---

```{r}
library(tidyverse)
```

# Core

## ggplo2

http://ggplot2.tidyverse.org

```{r}
ggplot(mpg, aes(displ, hwy, colour = class)) + geom_point()
```

## dplyr

http://dplyr.tidyverse.org

```{r}
starwars %>% 
  filter(species == "Droid")
```


## tidyr

http://tidyr.tidyverse.org

gather and spread
separate extract

## readr

http://readr.tidyverse.org

read files; faster; consistent function naming conventions

## purr

http://readr.tidyverse.org

functional programming

map
map_dbl

## tibble

http://tibble.tidyverse.org

do less and complain more
create tibble row-by-row with tribble

```{r}
tribble(
  ~x, ~y,  ~z,
  "a", 2,  3.6,
  "b", 1,  8.5
)
```


# Non-Core

## Import

for non-flat files/foreign files

### readxl

for xls and xlsx

### haven

for SPSS, stata, and SAS

### jsonlite

https://github.com/jeroen/jsonlite#jsonlite
parse json

### xml2
https://github.com/r-lib/xml2
parse xml

### httr
https://github.com/r-lib/httr
wrapper for curl package

### rvest

https://github.com/hadley/rvest
scraping web pages

### DBI
https://github.com/r-dbi/DBI
defines common interface between R and RDBMS (RSQLite, RPostgres, odbc)

## Wranlge

### stringr
http://stringr.tidyverse.org
for string manipulation


### lubridate
http://lubridate.tidyverse.org
date time manipulation

### forcats
http://forcats.tidyverse.org
handling categorical vairables

### hms
https://github.com/tidyverse/hms
for dealing with times

### blob
https://github.com/tidyverse/blob
binary vectors

## Program

### rlnag
http://rlang.tidyverse.org
tidyeval framework; quasiquotation and quosures

### magrittr
http://magrittr.tidyverse.org
the pipe
"%<>%"


### glue
https://github.com/tidyverse/glue
better version of paste

## Model

### recipes
https://github.com/topepo/recipes
for making design matrices


### rsample
https://github.com/topepo/rsample
bootstrapping etc.

### modelr
https://github.com/tidyverse/modelr
integrate models with tidyverse 

### R4DS
http://r4ds.had.co.nz
R for Data Science

### broom
https://github.com/tidyverse/broom

tidy, augment, glance