---
title: "R Notebook"
output: html_notebook
---

Get Data

```{r}
library(tidyverse)
library(caret)
```


```{r}
con = dbConnect(pg, user="kthomas1", password="",
                host="localhost", port=5432, dbname="bball")

raw_games <- dbGetQuery(con, "select * from oct_games") %>% 
    mutate(home = case_when(location == "Home" ~ 1, 
                            location == "Away" ~ 0)) %>% 
    select(-team_name, -location)

raw_schools <- dbGetQuery(con, "select * from oct_schools") %>% 
    select(-year) %>% 
    filter(division_id == 1)

raw_kp_def <- dbGetQuery(con, "select * from kp_def") %>% 
    select(-team_name, -TeamName, -starts_with("Rank")) %>% 
    rename(year = Season) %>% 
    rename_at(vars(-team_id, -year), funs(paste0("d_", .)))

raw_kp_off <- dbGetQuery(con, "select * from kp_off") %>% 
    select(-team_name, -TeamName) %>% 
    rename(year = Season) %>% 
    rename_at(vars(-team_id, -year), funs(paste0("o_", .)))

raw_kp_phys <- dbGetQuery(con, "select * from kp_phys") %>% 
    select(-team_name, -TeamName) %>% 
    rename(year = Season) %>% 
    select(year, team_id, Size, SizeRank, Exp, ExpRank, HgtEff, HgtEffRank, Bench, BenchRank)

raw_kp_points <- dbGetQuery(con, "select * from kp_points") %>% 
    select(-team_name, -TeamName) %>% 
    rename(year = Season)

raw_kp_misc <- dbGetQuery(con, "select * from kp_misc") %>% 
    select(-team_name, -TeamName) %>% 
    rename(year = Season) %>% 
    select(-contains("Opp"), -FG3Rate, -RankFG3Rate, -DFP)

```

geo mapping
notable statistics


Base Line coin flip

```{r}
coin <- raw_games %>% 
    inner_join(s)
    select(win)
```




Home Game
Clyde
Opionated Logistic Regression
Ridge Logistic Regression
Lasso Logistic Regression
Elastic Net Logistic Regression
PCA Regression
SVM
PageRank

# Feature Matrix

```{r}
wins <- raw_games %>% 
    select(team_id, game_date, win) %>% 
    mutate(game_date = as.Date(game_date, format = "%m/%d/%Y"))

#4 features
features_school <- raw_schools %>% 
    select(team_id, games, wins, losses, total_winpct)

#47 features
features_school_annual <- raw_kp_def %>% 
    inner_join(raw_kp_off) %>% 
    inner_join(raw_kp_points) %>% 
    inner_join(raw_kp_phys) %>% #currently just a handful of variables
    inner_join(raw_kp_misc) %>% 
    distinct(year, team_id, .keep_all = TRUE)

fgd <- raw_games %>% 
    inner_join(features_school) %>%  #will get us down to d-1 schools
    select(-roll_ot_year, -roll_ot_total) %>% 
    select(year, game_date, team_id, opponent_id, starts_with("roll"), -contains("total")) %>% 
    mutate(game_date = as.Date(game_date, format = "%m/%d/%Y")) 

fgd_both <- fgd %>% 
    inner_join(fgd, by = c("game_date" = "game_date", "opponent_id" = "team_id"))

fgdx <- fg_both %>% select(contains(".x")) %>% select(-year.x) 
fgdy <- fg_both %>% select(contains(".y")) %>% select(-year.y, -opponent_id.y) 
fgdz <- cbind(fgd_both[1:4], fgdx - fgdy)

fga <- raw_games %>% 
    inner_join(features_school) %>%  #will get us down to d-1 schools
    select(year, game_date, team_id, opponent_id) %>% 
    mutate(game_date = as.Date(game_date, format = "%m/%d/%Y")) %>% 
    inner_join(features_school_annual) %>% 
    arrange(team_id, year)

fga_both <- fga %>% 
    inner_join(features_school_annual, by = c("year" = "year", "opponent_id" = "team_id"))

fgax <- fga_both %>% select(contains(".x")) 
fgay <- fga_both %>% select(contains(".y")) 
fgaz <- cbind(fga_both[1:4], fgax - fgay)

features <- inner_join(fgdz, fgaz) %>% 
    inner_join(wins) %>% 
    select(win, contains(".x")) %>% 
    select(-year.x)

features[is.na(features)] <- 0

ws <- features$win
features <- as.data.frame(scale(features[,2:ncol(features)]))

```


```{r}

```


?? for below
```{r}
library(elasticnet)
x <- as.matrix(features)
object <- enet(x = x, y = ws)
fits <- predict.enet(object, x, type="fit")
preds <- fits$fit

library(glmnet)
gln <- glmnet(x, ws, family="binomial", alpha = 0.5, lambda = 0.001)
preds <- predict(gln, x, type = "response")
predz <- as.data.frame(preds) %>% mutate(winz = if_else(s0>.5,1,0))
table(ws, predz$winz)
```

above has 77% hit rate

problems: data is not in time; phys is fine; but other data points need to be on a rolling year basis



# Home Game

```{r}
clyde <- features %>% 
    select(team_id, year, game_date, win, roll_pscore_year, roll_pallow_year, roll_games_year, roll_pscore_year_opp, roll_pallow_year_opp, roll_games_year_opp) %>% 
    group_by(team_id, year) %>% 
    mutate(px = (lag(roll_pscore_year) + lag(roll_pallow_year_opp))/(2 * lag(roll_games_year)),
           py = (lag(roll_pscore_year_opp) + lag(roll_pallow_year))/(2 * lag(roll_games_year_opp)),
           clyde_win = if_else(px > py, 1, 0))
```



```{r}
confusionMatrix(table(clyde$win, clyde$clyde_win))
```

Note: if you don't lag the values you get a very positive 65%

cross validate clyde? roc curve?



```{r}
library(glmnet)
data(iris)
x <- model.matrix(Sepal.Length~., iris)[,-1]
y <- iris$Sepal.Length
m <- cv.glmnet(x, y)

fit = glmnet(x, features_y$win, family = "binomial")
```

```{r}
library(elasticnet)
enet(x = xx, y = features_y, lambda = 0)

xx <- features %>% select(win, Size, total_winpct)
fit <- glm(win ~ Size + total_winpct, data = features, family = "binomial")
probs <- predict(fit, type = "response")
pred <- rep(0, 128344)
pred[probs>.5] <- 1
table(features$win, pred)
```




probs determiend by 




Old Code

```{r}
#old seed
seeds <- data.frame(c(1,16,8,9,5,12,4,13,6,11,3,14,7,10,2,15))
colnames(seeds) <- "seed"

seeds_x <- seeds %>% 
    mutate(team2 = lead(seed)) %>% 
    mutate(row = row_number()) %>% 
    mutate(drp = if_else(row %% 2 == 0, "drop", "not")) %>% 
    filter(drp=="not") %>% 
    select(seed, team2) %>% 
    rename(team1 = seed)

round <- seeds_x %>% 
    mutate(win = if_else(team1 < team2, "win", "lose")) %>% 
    filter(win == "win")


```

```{r}
#conference rankings?
conf <- raw_schools %>% 
    select(team_id, conference_key)

conf_off <- raw_kp_off %>% 
    inner_join(conf) %>% 
    group_by(conference_key) %>% 
    summarise(ave1 = mean(RankeFG_Pct),
              ave2 = mean(RankTO_Pct),
              ave3 = mean(RankOR_Pct),
              ave4 = mean(RankFT_Rate)) %>% 
    mutate(ave_off = rowMeans(ave1 + ave2))

```

